---
title: "Validation of Nivolumab Model and Bayes Estimation"
author: "Jim Hughes"
date: "3 April 2019"
output:
  word_document: default
  html_document: default
---
### Introduction
To use the `mrgsolve` model and Bayes function for this study, they will need to
be validated against the gold standard, e.g. NONMEM.
```{r setup, include=FALSE}
# Prepare workspace
# Load package libraries
  library(dplyr)	# dplyr required for mrgsolve
  library(mrgsolve)	  # Metrum differential equation solver for pharmacometrics
  library(ggplot2)  # graphical package
  # library(MASS)  # mvrnorm in trunc_mvrnorm
  # library(MBESS)  # cor2cov in trunc_mvrnorm
  
# Set ggplot2 theme
  theme_bw2 <- theme_set(theme_bw(base_size = 14))
  theme_update(plot.title = element_text(hjust = 0.5))
  
# Set colourblind palette
  cbPalette <- data.frame(
		grey = "#999999",
		orange = "#E69F00",
		skyblue = "#56B4E9",
		green = "#009E73",
		yellow = "#F0E442",
		blue = "#0072B2",
		red = "#D55E00",
		pink = "#CC79A7",
		stringsAsFactors = F
  )
  
# Set master directory
  master_dir <- "C:/Users/hugjh001/Documents/nivo_sim/"
  
# Source external scripts
  script_path <- paste0(master_dir, "scripts/model_bayes_validation/")
  source(paste0(master_dir, "scripts/functions_utility.R"))  # functions utility

# - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
# Validation of mrgsolve model
# 1. Source model
# New objects: mod
  source(paste0(master_dir, "models/NivoPKTS_Final.R"))  # PopPK model script 
  
# 2. Create study population
# New objects: pop_df
  source(paste0(script_path, "create_population.R"))

# 3. Create input data set for mrgsolve and NONMEM simulation
# New objects: input_mrgsim_df input_nonmem_df output_mrgsim_df
  source(paste0(script_path, "create_sim_input.R"))
  
# 4. Run NONMEM simulation externally
  
# 5. Process mrgsolve and NONMEM output and validate mrgsolve model
# New objects: output_nonmem_df p_CvT p_TvT
  source(paste0(master_dir, "nonmem/processSIMdata.R"))
  run.name <- "NivoPKTS_NM_Sim"
  output_nonmem_df <- read.csv(
      paste0(master_dir, "output/", run.name, ".nm7/", run.name, ".fit.csv"),
      stringsAsFactors = F, na.strings = "."
    ) %>% tibble::as_tibble() %>%
    filter(AMT == 0)
  source(paste0(script_path, "process_sim_output.R"))
```
### Aim
1. Validate `mrgsolve` model against NONMEM model
2. Validate R Bayes function against NONMEM Bayes estimation
    + Sensitivity Analysis of ETAs to estimate
    + Testing estimation of different numbers of ETAs
    
### 1. Model Validation
Simulated model using a population of 100 individuals. Model used population as
described in `nivo_population.rmd`. This population was prepared for use with
both NONMEM and `mrgsolve`.

All validation plots below have the NONMEM predictions in blue, with the line
representing the median of the dependent variable, while the bands represent the
upper and lower 95% confidence intervals.

#### Concentration vs. Time

```{r model_plots, echo = F}
  p_CvT
  p_TvT
```

These plots show that the `mrgsolve` model predictions are equivalent to those
from the NONMEM model.

### 2. Bayes Validation
While Bayes estimation could be done with all of the ETAs in the model, it 
should be recognised that Bayes estimates are based on trough concentrations. It
is unclear how sensitive trough concentrations are to changes in each of the 
parameters in the model.

#### Sensitivity Analysis
To determine how sensitive the trough concentrations are to each ETA, the model
was simulated with ETA values of -1.2, -0.8, -0.4, 0, 0.4, 0.8 and 1.2 at two 
different time points: 14, 112 and 364 days. This was to capture the influence of 
the variability throughout the range of ETAs while also capturing the time
dependence of clearance and tumour size.

```{r sensitivity, echo = F}
# Source sensitivity analysis script
  source(paste0(master_dir, "scripts/190403_Sensitivity_Analysis.R"))

# Present results in a table
  output_sa_df %>%
    dplyr::mutate(ETA = c(
      "ETA_CL", "ETA_V1", "ETA_V2", "ETA_EMAX", 
      "ETA_EC50", "ETA_TG", "ETA_R", "ETA_IIVHAZ", "ETA_TMAX"
    )) %>%
    knitr::kable()
```

This table shows the sensitivity index $SI$ for each value. It is calculated
such that if the ETA is changed by $\delta ETA$ then the trough concentration is
changed by $\delta ETA * SI$ on average.

This shows that clearance is the most influential parameter. TMAX (the parameter
responsible for time-varying clearance) is also very influential, but only at
later time-points. It appears that EMAX, EC50, TG and R are also time-dependent.

As changes in the EC50, R and IIVHAZ do not have an absolute SI over 1 at 364 days,
these are candidates for removal. For Bayes estimation of early time points, 
where EMAX, TG and TMAX are less influential they may need to be removed as well.

#### Validation of Different ETA models
The candidate models for validation estimate ETAs as following:

* Full - All ETAs (no IIVHAZ) [8 ETAs]
* Base - All "influential" ETAs (no EC50, R or IIVHAZ) [6 ETAs]
* NoTG - Base with no TG ETA [5 ETAs] - done
* NoEMAX - Base with no EMAX ETA [5 ETAs]
* NoTMAX - Base with no TMAX ETA [5 ETAs]
* k

```{r valload, include = F}
# Source validation script
  source(paste0(master_dir, "scripts/190404_Bayes_Validation.R"))
```

```{r valplot, echo = F, fig.height = 5.6, fig.width = 8.5}
  names(val_plot_lst) <- c("Full", "Base", "NoTG", "NoEMAX", "NoTMAX")
  val_plot_lst
```

Plots show the absolute error of the R Bayes method with respect to NONMEM. This
means that the absolute error for the ETAs from both methods (R Bayes and NONMEM)
were calculated, and then the difference between those two values was found.

Points on plots show the median error, while the error bars show the
interquartile range.

These plots show that overall the R Bayes function performs less favourably
than NONMEM, but still manage to keep the line of identity with the 
interquartile range.

The most influential parameters are reasonalby well estimated (being CL and Tmax).
The R Bayes method had difficulty estimating Emax and V2, which are of
moderate influence.