---
title: "Building Nivolumab Population"
author: "Jim Hughes"
date: "15 March 2019"
output:
  word_document: default
  html_document: default
---
### Aim
Create a population that is representative of the data used to develop the
Nivolumab models shown in:

> Bajaj G, Wang X, Agrawal S, Gupta M, Roy A, Feng Y. Model-Based Population
Pharmacokinetic Analysis of Nivolumab in Patients With Solid Tumors. CPT
Pharmacometrics Syst Pharmacol. 2017;6(1):58-66.

> Liu C, Yu J, Li H, Liu J, Xu Y, Song P, et al. Association of time-varying
clearance of nivolumab with disease dynamics and its implications on exposure
response analysis. Clin Pharmacol Ther. 2017;101(5):657-66.

This population is required to meet the following criteria:

* Must include weight, height, age, sex, ECOG, serum creatinine and albumin
* Albumin variability representative of metastatic melanoma patients
* Weight and height are appropriately correlated
* eGFR as determined by CKD-EPI representative of metastatic melanoma patients
* Cancer type as metastatic melanoma
* Unknown ADA assay results

### Demographic Data (Bajaj)
The demographic data from Bajaj et al. states the means and standard deviation.
These values are:

* Age - 61.12 years (11.12) [23 - 87]
* Weight - 79.09 kg (19.28) [34 - 162]
* eGFR - 78.49 mL/min/1.73m^2 (21.63)
* Sex - 66.7% male, 33.3% female
* ECOG - 38.73% 0, 61.26% >0

While it is possible to create a population using eGFR as is, this will ignore
the it's correlation with SeCr, gender and age. While the units for eGFR look
like they are normalised for body surface area, this is actually the units 
provided by the CKD-EDI equaion. Therefore it is completely independent of 
weight (and height).

### Demographic Data (Not Bajaj)
#### eGFR
When observing how other papers have handled eGFR, there are mixed levels of 
detail. Two influential papers from the Journal of Clinical Oncology show that 
they used covariate data from real patients sourced from clinical trial data. 
Reference below:

> Zhao X, Suryawanshi S, Hruska M, Feng Y, Wang X, Shen J, et al. Assessment of
nivolumab benefit-risk profile of a 240-mg flat dose relative to a 3-mg/kg
dosing regimen in patients with advanced tumors. Ann Oncol. 2017;28(8):2002-8.

> Long GV, Tykodi SS, Schneider JG, Garbe C, Gravis G, Rashford M, et al.
Assessment of nivolumab exposure and clinical safety of 480 mg every 4 weeks
flat-dosing schedule in patients with cancer. Annals of oncology : official
journal of the European Society for Medical Oncology. 2018;29(11):2208-13.

A simulation study looking at both pembrolizumab and nivolumab randomly sampled
from normal distributions to obtain their covariates. Unfortunately they 
sampled eGFR values directly from a normal distribution and did not mention any
correlation between covariate values. Reference below:

> Ogungbenro K, Patel A, Duncombe R, Nuttall R, Clark J, Lorigan P. Dose
Rationalization of Pembrolizumab and Nivolumab Using Pharmacokinetic Modeling
and Simulation and Cost Analysis. Clin Pharmacol Ther. 2018;103(4):582-90.

#### Albumin
Datta M et al. discuss serum calcium, albumin and tumour stage in malignant
melanomas in their paper. According to this paper metastatic melanoma is 
associated with a "statistically significant" (p < 0.05) decrease in albumin. 
This paper states the mean and standard deviation as:

* Stage III - 4.15 g/dl (0.33) [2.5 - 5.1]
* Stage IV - 3.92 g/dl (0.45)

These values match with the average value used in the model by Liu et al. of
4 mg/dL, but not the units. As 4 mg/dL is 1000x lower than the normal value
it is assumed that the units used by Liu et al. are a typographical error. 
It is assumed that albumin would be normally distributed. Reference below:

> Datta M, Savage P, Lovato J, Schwartz GG. Serum calcium, albumin and tumor 
stage in cutaneous malignant melanoma. Future Oncol. 2016;12(19):2205-14.

### Covariate Sampling
Weight, height, age, sex, ECOG, serum creatinine and albumin are required to 
simulate using the Liu et al. model. The distributions taken straight from
the literature are:

* Age - 61.12 years (11.12) [23 - 87]
* Weight - 79.09 kg (19.28) [34 - 168]
* Sex - 66.7% male, 33.3% female
* ECOG - 38.73% 0, 61.26% >0
* Albumin - 3.92 (0.45) [2.5 - 5.1]

Serum creatinine were determined from external literature as stated above:

* Serum Creatinine - 78.3 (20) [50, 200]

Serum creatinine were based on these distributions found in literature.

* eGFR - 78.49 mL/min/1.73m^2 (21.63)

Weight and height will be correlated with a Pearsons correlation coefficient of 
0.7. Age and serum creatinine _could_ be correlated, however it is not clear 
from literature what the Pearsons correlation would be, references have shown
values ranging from 0.2 to 0.8. 0.2 seems more reasonable...

```{r setup, include = F}
# Prepare work environment
# Load libraries
  library(MASS)
  library(MBESS)
  library(ggplot2)
  library(cowplot)

# Define colourblind palette and custom palette
  cbPalette <- data.frame(
    grey = "#999999",
    orange = "#E69F00",
    skyblue = "#56B4E9",
    green = "#009E73",
    yellow = "#F0E442",
    blue = "#0072B2",
    red = "#D55E00",
    pink = "#CC79A7",
    stringsAsFactors = F
  )

# Set ggplot2 theme
  theme_bw2 <- theme_set(theme_bw(base_size = 14))
  theme_update(plot.title = element_text(hjust = 0.5))

# Load functions utility
  source("C:/Users/hugjh001/Documents/nivo_sim/scripts/functions_utility.R")

# Define demographic data
# Number of individuals
  nid <- 100000

# Continuous
  mean.AGE <- 61.12
  sd.AGE <- 11.12
  range.AGE <- c(23, 87)
  
  mean.ALB <- 3.92
  sd.ALB <- 0.45
  range.ALB <- c(2.5, 5.1)
  
  mean.SECR.M <- 90.6*1.04
  mean.SECR.F <- 71.5*1.04
  sd.SECR <- 28
  range.SECR.M <- c(0, 300)
  range.SECR.F <- c(0, 300)
  
  mean.WT <- 79.09
  sd.WT <- 19.28
  range.WT <- c(34, 168)
 
# Categorical
  sex.prob <- 0.667  # males
  ecog.prob <- 0.6126  # >0
 
# Comparators
  med.BMI <- 24.9
  range.BMI <- c(13.5, 46.6)
  cat.BMI <- data.frame(
   CAT = c("Underweight", "Normal", "Overweight", "Obese"),
   PER = c(4.1, 46.3, 38.6, 11)
  )
  
  mean.eGFR <- 78.49
  sd.eGFR <- 21.63
  range.eGFR <- c(0, 250)
 
# Sample covariate values from distributions
# Age & Albumin - normal distribution
  AGE <- trunc.rnorm(n = nid, mean = mean.AGE, sd = sd.AGE, range = range.AGE)
  ALB <- trunc.rnorm(n = nid, mean = mean.ALB, sd = sd.ALB, range = range.ALB)
  
# Sex & ECOG - binomial distribution
  SEX <- rbinom(nid, 1, sex.prob)
  ECOG <- rbinom(nid, 1, ecog.prob)
 
# Weight & Serum Creatinine - log-normal distribution
  WT <- trunc.rnorm(n = nid, mean = mean.WT, sd = sd.WT, 
    range = range.WT, log = T)
  SECR.M <- trunc.rnorm(n = length(SEX[SEX == 1]), mean = mean.SECR.M, 
    sd = sd.SECR, range = range.SECR.M, log = T)
  SECR.F <- trunc.rnorm(n = length(SEX[SEX == 0]), mean = mean.SECR.F, 
    sd = sd.SECR, range = range.SECR.F, log = T)
  
# Bind covariate data.frame
  cov.df <- data.frame(AGE, ALB, WT, SEX, ECOG)
  cov.df$SECR[cov.df$SEX == 1] <- SECR.M
  cov.df$SECR[cov.df$SEX == 0] <- SECR.F
 
# Calculated eGFR
  cov.df$cGFR <- apply(cov.df, 1, function(df) {
    ckdepi.fn(df["SECR"], df["AGE"], df["SEX"], 0)
  })

# Randomly Sampled GFR correlated to AGE
  mean.AGE.GFR <- c(mean.AGE, mean.eGFR)
  sd.AGE.GFR <- c(sd.AGE, sd.eGFR)
  
  corr.AGE.GFR <- matrix(c(
     1.00, -0.34,
    -0.34,  1.00
  ), nrow = 2, ncol = 2)  # symmetrical correlation matrix
  
  lower.AGE.GFR <- c(range.AGE[1], range.eGFR[1])
  upper.AGE.GFR <- c(range.AGE[2], range.eGFR[2])
  
  corrsim <- trunc.mvrnorm(n = nid, mean = mean.AGE.GFR, sd = sd.AGE.GFR, 
    corr_mat = corr.AGE.GFR, lower = lower.AGE.GFR, upper = upper.AGE.GFR, 
    log = F)
  cov.df$mvAGE <- corrsim[, 1]
  cov.df$mvGFR <- corrsim[, 2]
  
# Add SEXf column
  cov.df$SEXf <- factor(cov.df$SEX)
  levels(cov.df$SEXf) <- c("Female", "Male")
```

### Visualise Distributions

```{r covplots, echo = F, fig.asp = 1.5}
  p <- NULL
  p <- ggplot(data = cov.df)
  plot.AGE <- p + geom_histogram(aes(AGE), bins = 30, fill = cbPalette$blue)
  plot.AGE <- plot.AGE + geom_vline(xintercept = range.AGE, linetype = "dashed")
  
  plot.ALB <- p + geom_histogram(aes(ALB), bins = 30, fill = cbPalette$orange)
  plot.ALB <- plot.ALB + geom_vline(xintercept = range.ALB, linetype = "dashed")
  
  plot.SECR <- p + geom_histogram(aes(SECR), bins = 30, fill = cbPalette$pink)
  plot.SECR <- plot.SECR + geom_vline(xintercept = range.SECR.F, linetype = "dashed")
  
  plot.WT <- p + geom_histogram(aes(WT), bins = 30, fill = cbPalette$green)
  plot.WT <- plot.WT + geom_vline(xintercept = range.WT, linetype = "dashed")
  
  plot.hist <- plot_grid(plot.AGE, plot.ALB, plot.SECR, plot.WT, 
    ncol = 2)
  plot.hist
```

This plot shows the truncated distributions for each of the continuous 
covariates used in the nivolumab population. Age and albumin are normally
distributed, weight, height and serum creatinine are log-normally distributed.

```{r gfrplots, echo = F}
  plot.GFR <- p + geom_histogram(aes(cGFR, fill = cbPalette$red), 
    bins = 30, alpha = 0.5)
  plot.GFR <- plot.GFR + geom_histogram(aes(mvGFR, fill = cbPalette$green), 
    bins = 30, alpha = 0.5)
  plot.GFR <- plot.GFR + scale_fill_identity(name = "Dist", guide = "legend", 
    labels = c("Sampled GFR", "Calculated GFR"))
  plot.GFR <- plot.GFR + xlab("eGFR (mL/min/1.73m2)")
  plot.GFR
```

While the calculated eGFR using CKD-EPI looks like a strange distribution, this 
is the distribution that GFR would have given a set of log-normally distributed
serum creatinine values.

```{r agegfrplots, echo = F}
  plot.AGE.GFR <- p + geom_point(aes(x = AGE, y = cGFR), 
    colour = cbPalette$red, shape = 1, alpha = 0.1)
  plot.AGE.GFR <- plot.AGE.GFR + xlab("Age (years)")
  plot.AGE.GFR <- plot.AGE.GFR + ylab("eGFR (mL/min/1.73m2)")
  plot.AGE.GFR <- plot.AGE.GFR + coord_cartesian(xlim = c(23, 87), 
    ylim = c(0, 250))

  plot.mvAGE.mvGFR <- p + geom_point(aes(x = mvAGE, y = mvGFR), 
    colour = cbPalette$green, shape = 1, alpha = 0.1)
  plot.mvAGE.mvGFR <- plot.mvAGE.mvGFR + xlab("Age (years)")
  plot.mvAGE.mvGFR <- plot.mvAGE.mvGFR + ylab("eGFR (mL/min/1.73m2)")
  plot.mvAGE.mvGFR <- plot.mvAGE.mvGFR + coord_cartesian(xlim = c(23, 87), 
    ylim = c(0, 250))
  
  plot.AGE.eGFR <- plot_grid(plot.AGE.GFR, plot.mvAGE.mvGFR)
  plot.AGE.eGFR
```

The CKD-EPI equation enforces a harder correlation between age and eGFR at 
higher GFR values, while has a softer correlation at lower eGFR values.

```{r secrgfrplots, echo = F}
  plot.SECR.GFR <- p + geom_point(aes(x = SECR, y = cGFR, colour = SEXf), 
    shape = 1, alpha = 0.1)
  plot.SECR.GFR <- plot.SECR.GFR + xlab("Serum Creatinine (umol/L)")
  plot.SECR.GFR <- plot.SECR.GFR + ylab("eGFR (mL/min/1.73m2)")
  plot.SECR.GFR <- plot.SECR.GFR + guides(
    colour = guide_legend(override.aes = list(alpha=1))
  )
  plot.SECR.GFR
```

Here the two distinct distribution curves can be seen, the main one (and the
lower peak of the histogram) can be seen on the right. While the minor one
(and the taller peak) can be seen on the left. Of course males and females 
have different Serum creatinine due to the correlation of serum creatinine with
sex during simulation.