---
title: "Building Nivolumab Population"
author: "Jim Hughes"
date: "15 March 2019"
output:
  word_document: default
  html_document: default
---
### Aim
Create a population that is representative of the data used to develop the
Nivolumab models shown in:

> Bajaj G, Wang X, Agrawal S, Gupta M, Roy A, Feng Y. Model-Based Population
Pharmacokinetic Analysis of Nivolumab in Patients With Solid Tumors. CPT
Pharmacometrics Syst Pharmacol. 2017;6(1):58-66.

> Liu C, Yu J, Li H, Liu J, Xu Y, Song P, et al. Association of time-varying
clearance of nivolumab with disease dynamics and its implications on exposure
response analysis. Clin Pharmacol Ther. 2017;101(5):657-66.

This population is required to meet the following criteria:

* Must include weight, height, age, sex, ECOG, serum creatinine and albumin
* Albumin variability representative of metastatic melanoma patients
* Weight and height are appropriately correlated
* eGFR as determined by CKD-EPI representative of metastatic melanoma patients
* Cancer type as metastatic melanoma
* Unknown ADA assay results

### Demographic Data (Bajaj)
The demographic data from Bajaj et al. states the means and standard deviation.
These values are:

* Age - 61.12 years (11.12) [23 - 87]
* Weight - 79.09 kg (19.28) [34 - 162]
* eGFR - 78.49 mL/min/1.73m^2 (21.63)
* Sex - 66.7% male, 33.3% female
* ECOG - 38.73% 0, 61.26% >0

While it is possible to create a population using eGFR as is, this will ignore
the it's correlation with SeCr, gender and age. Also the units for eGFR as used
in the model are normalised for body surface area. This means that weight and 
height are also correlated with this value. 

### Demographic Data (Not Bajaj)
#### eGFR
When observing how other papers have handled the dependence of eGFR on knowing 
body surface area while simulating, there are mixed levels of detail. Two 
influential papers from the Journal of Clinical Oncology show that they used 
covariate data from real patients sourced from clinical trial data. Reference 
below:

> Zhao X, Suryawanshi S, Hruska M, Feng Y, Wang X, Shen J, et al. Assessment of
nivolumab benefit-risk profile of a 240-mg flat dose relative to a 3-mg/kg
dosing regimen in patients with advanced tumors. Ann Oncol. 2017;28(8):2002-8.

> Long GV, Tykodi SS, Schneider JG, Garbe C, Gravis G, Rashford M, et al.
Assessment of nivolumab exposure and clinical safety of 480 mg every 4 weeks
flat-dosing schedule in patients with cancer. Annals of oncology : official
journal of the European Society for Medical Oncology. 2018;29(11):2208-13.

A simulation study looking at both pembrolizumab and nivolumab randomly sampled
from normal distributions to obtain their covariates. Unfortunately they 
sampled eGFR values directly from a normal distribution and did not mention any
correlation between covariate values. Reference below:

> Ogungbenro K, Patel A, Duncombe R, Nuttall R, Clark J, Lorigan P. Dose
Rationalization of Pembrolizumab and Nivolumab Using Pharmacokinetic Modeling
and Simulation and Cost Analysis. Clin Pharmacol Ther. 2018;103(4):582-90.

Due to a lack of guidance from these papers, it was decided to source BMI, BSA
or height distributions from other papers involving a representative population.

#### BMI
Cortellini et al. explored the relationship between BMI and the outcome of 
therapy with checkpoint inhibitors. Their analysis was done using a population
of 976 cancer patients. 18.7% of these patients were diagnosed with melanoma,
while 72.3% of patients were taking nivolumab. 20% of the patients would be
representative. Medians, ranges and proportions of demographics and reference 
below:

* Age - 68 years [24 - 92]
* Weight - 71 kg [35 - 139]
* Sex - 67.9% male, 32.1% female
* BMI - 24.9 [13.5 - 46.6]
* BMI Category - Underweight 4.1%, Normal 46.3%, Overweight 38.6%, Obese 11%

> Cortellini A, Bersanelli M, Buti S, Cannita K, Santini D, Perrone F, et al. 
A multicenter study of body mass index in cancer patients treated with 
anti-PD-1/PD-L1 immune checkpoint inhibitors: When overweight becomes favorable. 
Journal for ImmunoTherapy of Cancer. 2019;7(1).

Height can be sampled in such a way that it matches the BMI values here. This 
can then be used to determine eGFR using a normally distributed serum 
creatinine, that is based on the eGFR distribution from the paper by Bajaj et 
al.

#### Albumin
Datta M et al. discuss serum calcium, albumin and tumour stage in malignant
melanomas in their paper. According to this paper metastatic melanoma is 
associated with a "statistically significant" (p < 0.05) decrease in albumin. 
This paper states the mean and standard deviation as:

* Stage III - 4.15 g/dl (0.33) [2.5 - 5.1]
* Stage IV - 3.92 g/dl (0.45)

These values match with the average value used in the model by Liu et al. of
4 mg/dL, but not the units. As 4 mg/dL is 1000x lower than the normal value
it is assumed that the units used by Liu et al. are a typographical error. 
It is assumed that albumin would be normally distributed. Reference below:

> Datta M, Savage P, Lovato J, Schwartz GG. Serum calcium, albumin and tumor 
stage in cutaneous malignant melanoma. Future Oncol. 2016;12(19):2205-14.

### Covariate Sampling
Weight, height, age, sex, ECOG, serum creatinine and albumin are required to 
simulate using the Liu et al. model. The distributions taken straight from
the literature are:

* Age - 61.12 years (11.12) [23 - 87]
* Weight - 79.09 kg (19.28) [34 - 168]
* Sex - 66.7% male, 33.3% female
* ECOG - 38.73% 0, 61.26% >0
* Albumin - 3.92 (0.45) [2.5 - 5.1]

Height and serum creatinine were determined from external literature as stated 
above:

* Height - 168.86 (10) [150 - 195]
* Serum Creatinine - 78.3 (20) [50, 200]

Both height and serum creatinine were based on these distributions
found in literature.

* eGFR - 78.49 mL/min/1.73m^2 (21.63)
* BMI - 24.9 [13.5 - 46.6]
* BMI Category - Underweight 4.1%, Normal 46.3%, Overweight 38.6%, Obese 11%

Weight and height will be correlated with a Pearsons correlation coefficient of 
0.7. Age and serum creatinine _could_ be correlated, however it is not clear 
from literature what the Pearsons correlation would be, references have shown
values ranging from 0.2 to 0.8. 0.2 seems more reasonable...

```{r setup, include = F}
# Prepare work environment
# Load libraries
  library(MASS)
  library(MBESS)
  library(ggplot2)
  library(cowplot)

# Define colourblind palette and custom palette
  cbPalette <- data.frame(
    grey = "#999999",
    orange = "#E69F00",
    skyblue = "#56B4E9",
    green = "#009E73",
    yellow = "#F0E442",
    blue = "#0072B2",
    red = "#D55E00",
    pink = "#CC79A7",
    stringsAsFactors = F
  )

# Set ggplot2 theme
  theme_bw2 <- theme_set(theme_bw(base_size = 14))
  theme_update(plot.title = element_text(hjust = 0.5))

# Load functions utility
  source("C:/Users/hugjh001/Documents/nivo_sim/scripts/functions_utility.R")

# Define demographic data
# Number of individuals
  nid <- 100000

# Continuous
  mean.AGE <- 61.12
  sd.AGE <- 11.12
  range.AGE <- c(23, 87)
  
  mean.ALB <- 3.92
  sd.ALB <- 0.45
  range.ALB <- c(2.5, 5.1)
  
  mean.HT <- 168.86
  sd.HT <- 10
  range.HT <- c(150, 200)
  
  mean.SECR <- 78.3
  sd.SECR <- 30
  range.SECR <- c(50, 200)
  
  mean.WT <- 79.09
  sd.WT <- 19.28
  range.WT <- c(34, 168)
 
# Categorical
  sex.prob <- 0.667  # males
  ecog.prob <- 0.6126  # >0
 
# Comparators
  med.BMI <- 24.9
  range.BMI <- c(13.5, 46.6)
  cat.BMI <- data.frame(
   CAT = c("Underweight", "Normal", "Overweight", "Obese"),
   PER = c(4.1, 46.3, 38.6, 11)
  )
  
  mean.eGFR <- 78.49
  sd.eGFR <- 21.63
 
# Sample covariate values from distributions
# Age & Albumin - normal distribution
  AGE <- trunc.rnorm(n = nid, mean = mean.AGE, sd = sd.AGE, range = range.AGE)
  ALB <- trunc.rnorm(n = nid, mean = mean.ALB, sd = sd.ALB, range = range.ALB)
 
# Weight and Height - multivariate log-normal distribution
  mean.WT.HT <- c(mean.WT, mean.HT)
  sd.WT.HT <- c(sd.WT, sd.HT)
  lower.WT.HT <- c(range.WT[1], range.HT[1])
  upper.WT.HT <- c(range.WT[2], range.HT[2])
  
# Set up correlation matrix for height and weight
  corr.WT.HT <- matrix(c(
    1.0, 0.7,
    0.7, 1.0
  ), nrow = 2, ncol = 2)  # symmetrical correlation matrix
  
# Use trunc.mvrnorm function
  corrsim <- trunc.mvrnorm(n = nid, mean = mean.WT.HT, sd = sd.WT.HT, 
    corr_mat = corr.WT.HT, lower = lower.WT.HT, upper = upper.WT.HT, log = T)
  
  WT <- corrsim[,1]
  HT <- corrsim[,2]
 
# Serum Creatinine - log-normal distribution
  SECR <- trunc.rnorm(n = nid, mean = mean.SECR, sd = sd.SECR, 
    range = range.SECR, log = T)
 
# Sex & ECOG - binomial distribution
  SEX <- rbinom(nid, 1, sex.prob)
  ECOG <- rbinom(nid, 1, ecog.prob)
 
# Determine 
# eGFR
  cov.df <- data.frame(AGE, ALB, WT, HT, SECR, SEX, ECOG)
  cov.df$eGFR <- apply(cov.df, 1, function(df) {
    ckdepi.fn(df["SECR"], df["AGE"], df["SEX"], 0)*1.73/bsa.fn(df["WT"], df["HT"])
  })
  
  cov.df$GFR <- trunc.rnorm(n = nid, mean = mean.eGFR, sd = sd.eGFR, 
    range = c(0, 150))
```

### Visualise Distributions

```{r covplots, echo = F, fig.asp = 1.5}
  p <- NULL
  p <- ggplot(data = cov.df)
  plot.AGE <- p + geom_histogram(aes(AGE), bins = 30, fill = cbPalette$blue)
  plot.AGE <- plot.AGE + geom_vline(xintercept = range.AGE, linetype = "dashed")
  
  plot.ALB <- p + geom_histogram(aes(ALB), bins = 30, fill = cbPalette$orange)
  plot.ALB <- plot.ALB + geom_vline(xintercept = range.ALB, linetype = "dashed")
  
  plot.SECR <- p + geom_histogram(aes(SECR), bins = 30, fill = cbPalette$pink)
  plot.SECR <- plot.SECR + geom_vline(xintercept = range.SECR, linetype = "dashed")
  
  plot.WT <- p + geom_histogram(aes(WT), bins = 30, fill = cbPalette$green)
  plot.WT <- plot.WT + geom_vline(xintercept = range.WT, linetype = "dashed")
  
  plot.HT <- p + geom_histogram(aes(HT), bins = 30, fill = cbPalette$yellow)
  plot.HT <- plot.HT + geom_vline(xintercept = range.HT, linetype = "dashed")
  
  plot.WT.HT <- p + geom_point(aes(x = WT, y = HT), 
    colour = cbPalette$red, shape = 1, alpha = 0.1, size = 0.5)
  plot.WT.HT <- plot.WT.HT + geom_vline(xintercept = range.WT, linetype = "dashed")
  plot.WT.HT <- plot.WT.HT + geom_hline(yintercept = range.HT, linetype = "dashed")
  
  plot.hist <- plot_grid(plot.AGE, plot.ALB, plot.SECR, plot.WT, plot.HT, plot.WT.HT, 
    ncol = 2)
  plot.hist
```

This plot shows the truncated distributions for each of the continuous 
covariates used in the nivolumab population. Age and albumin are normally
distributed, weight, height and serum creatinine are log-normally distributed.

```{r gfrplots, echo = F}
  plot.GFR <- p + geom_histogram(aes(eGFR, fill = cbPalette$red), 
    bins = 30, alpha = 0.5)
  plot.GFR <- plot.GFR + geom_histogram(aes(GFR, fill = cbPalette$blue), 
    bins = 30, alpha = 0.5)
  plot.GFR <- plot.GFR + scale_fill_identity(name = "Dist", guide = "legend", 
    labels = c("Najaj GFR", "Sampled GFR"))
  plot.GFR <- plot.GFR + xlab("eGFR (mL/min/1.73m2)")
  plot.GFR
```

The two distributions do not exactly overlap which is potentially problematic. 
The issue is
that reducing the mean serum creatinine (and therefore increasing the eGFR), 
results in a shift of the mean eGFR upwards but a decrease in the standard
deviation (so the distribution becomes taller). 

I think this is due to the weird relationship between serum creatinine and 
GFR. If no lower bound is set to serum creatinine, this is the relationship
between it and GFR as determined by CKD-EPI.

```{r ckdepi, echo = F}
# Serum Creatinine - log-normal distribution
  test.SECR <- trunc.rnorm(n = nid, mean = mean.SECR, sd = sd.SECR, 
    range = c(0, 200), log = T)
 
# Determine 
# eGFR
  test.df <- data.frame(AGE, ALB, WT, HT, test.SECR, SEX, ECOG)
  test.df$eGFR <- apply(test.df, 1, function(df) {
    ckdepi.fn(df["test.SECR"], df["AGE"], df["SEX"], 0)*1.73/bsa.fn(df["WT"], df["HT"])
  })
  
  plot.SECR <- NULL
  plot.SECR <- ggplot(data = test.df)
  plot.SECR <- plot.SECR + geom_point(aes(x = test.SECR, y = eGFR), 
    colour = cbPalette$blue, size = 0.5, shape = 1, alpha = 0.5)
  plot.SECR <- plot.SECR + geom_vline(xintercept = 50, linetype = "da")
  plot.SECR <- plot.SECR + xlab("Serum Creatinine (umol/L)")
  plot.SECR <- plot.SECR + scale_y_continuous("eGFR (mL/min/1.73m2)", 
    breaks = 0:5*30)
  plot.SECR
```

Which to me is a bit confusing, maybe this would be fixed with additional
correlation between covariates? but it's hard to say...