---
title: "Building Nivolumab Population - Tumour Size"
author: "Jim Hughes"
date: "8 May 2019"
output:
  word_document: default
  html_document: default
---
### Aim
Create a population that is representative of the data used to develop the
Nivolumab models shown in:

> Bajaj G, Wang X, Agrawal S, Gupta M, Roy A, Feng Y. Model-Based Population
Pharmacokinetic Analysis of Nivolumab in Patients With Solid Tumors. CPT
Pharmacometrics Syst Pharmacol. 2017;6(1):58-66.

> Liu C, Yu J, Li H, Liu J, Xu Y, Song P, et al. Association of time-varying
clearance of nivolumab with disease dynamics and its implications on exposure
response analysis. Clin Pharmacol Ther. 2017;101(5):657-66.

This population is required to meet the following criteria:

* ~~Must include weight, age, sex, ECOG, serum creatinine and albumin~~
* ~~Albumin variability representative of metastatic melanoma patients~~
* ~~Gender and serum creatinine are appropriately correlated~~
* ~~eGFR as determined by CKD-EPI representative of metastatic melanoma patients~~
* ~~Cancer type as metastatic melanoma~~
* ~~Unknown ADA assay results~~
* __Baseline tumour size that correlates with ECOG__

### Tumour Size of Model Demographics
While the model from Liu et al. contains an average tumour size, the demographic
data from Bajaj et al. does not state tumour size. Therefore, external sources
are required to determine the standard deviation and range of baseline tumour 
size in melanoma patients.

It is also important that baseline tumour size correlates with ECOG, though
it is unclear by how much. The proportion of ECOG for the model demographics is 
provided in Bajaj et al.

* Baseline Tumour Size - 54.6 mm (?) [? - ?]
* ECOG - 61.26% > 0

### Tumour Size of External Demographics

> Joseph RW, Elassaiss-Schaap J, Kefford R, Hwu W-J, Wolchok JD, Joshua AM, et 
al. Baseline Tumor Size Is an Independent Prognostic Factor for Overall Survival
in Patients with Melanoma Treated with Pembrolizumab. Clin Cancer Res. 
2018;24(20):4960.

This paper features a significant amount of info on its population tumour size,
in a melanoma population. Interestingly the baseline tumour size follows more of
a log-normal distribution, or potentially even an inverse gamma or cauchy 
distribution. Additionally, it has a coefficient of 120%! The median is different
but it is hard to compare baseline tumour size between studies.

This also provides information on baseline tumour size stratified by ECOG! The 
difference in percentage above median tumour size for different ECOG values is
statistically different (p < 0.001). This gives an idea of how the correlation 
should be handled.

* Baseline Tumour Size - 102 mm (122 mm) [? - ?]
* ECOG - 34% > 0
* ECOG 0 Tumour Size > 102 mm - 42%
* ECOG 1 Tumour Size > 102 mm - 65%

> Chatterjee MS, Elassaiss-Schaap J, Lindauer A, Turner DC, Sostelly A, 
Freshwater T, et al. Population Pharmacokinetic/Pharmacodynamic Modeling of 
Tumor Size Dynamics in Pembrolizumab-Treated Advanced Melanoma. CPT 
Pharmacometrics Syst Pharmacol. 2017;6(1):29-39.

Another PK paper using melanoma patients shows a similar median value. The 
mixture model for this seems to split its tumour into two compartments, with 
different variability terms on each that are heavily correlated.

This seems overly complicated for our simulation, but it provides a range on
their demographic data for baseline tumour size.

* Baseline Tumour Size Demographic Data - 80.6 mm (?) [10 - 895]
* Baseline Tumour Size Model Shallow Compartment - (44.81)
* ECOG - 35.58% > 0

### Correlation

Generating random correlated data between a binary and a continuous variable
can be done in a few different ways, but an easy way is using a Gaussian copula.
The copula package is another easy way of achieving this, but doesn't work in 
R version 3.4.4 encouraging the use of a dependency free method.

```{r setup, include = F}
# Load packages
  library(ggplot2)

# Source functions utility
  source("C:/Users/hugjh001/Documents/nivo_sim/scripts/functions_utility.R")

# Set ggplot2 theme
  theme_bw2 <- theme_set(theme_bw(base_size = 14))
  theme_update(plot.title = element_text(hjust = 0.5))
  
# Define colourblind palette and custom palette
  cbPalette <- data.frame(
    grey = "#999999",
    orange = "#E69F00",
    skyblue = "#56B4E9",
    green = "#009E73",
    yellow = "#F0E442",
    blue = "#0072B2",
    red = "#D55E00",
    pink = "#CC79A7",
    stringsAsFactors = F
  )
```

```{r copula}
# Define the correlation coefficient
  r <- 0.33  # 0.55
# Define the correlation matrix and use Choleski decomposition
  sigma <- matrix(c(1, r, r, 1), ncol = 2)
  s <- chol(sigma)
  
# Set number of samples and create a correlated distribution function (u)
  n <- 10000
  z <- s%*%matrix(rnorm(n*2), nrow = 2)
  u <- pnorm(z)

# Simulate tumour size using log-normal distribution
  bts_mean <- 52.7
  bts_sd <- bts_mean*0.8
  bts_mu <- log(bts_mean^2/sqrt((bts_sd^2 + bts_mean^2)))
  bts_sig <- sqrt(log(bts_sd^2/bts_mean^2 + 1))
  bts <- exp(qnorm(u[1, ], bts_mu, bts_sig))
  
# ECOG is binomial, so very simple to sample
  ecog1 <- u[2, ] > 0.4  # provides probability of ecog > 1 of ~0.612
  
# Observe the data
  c(mean(bts), sd(bts), sum(ecog1)/n)
  c(sum(bts[ecog1] > median(bts))/sum(ecog1), sum(bts[!ecog1] > median(bts))/sum(!ecog1))
  c(
    min = min(bts), quantile(bts, prob = 0.05), quantile(bts, prob = 0.25), 
    quantile(bts, prob = 0.5), quantile(bts, prob = 0.75), 
    quantile(bts, prob = 0.95), max = max(bts)
  )
  
  plot(density(bts))

  plot(density(bts[!ecog1]), main = "bts by ecog1", xlab = "bts")
  lines(density(bts[ecog1]), lty = 2)
  legend('topright', c("ecog0", "ecog1"), lty = c(1,2))
  
  p <- NULL
  p <- ggplot(data = data.frame(logBTS = log(bts), ECOG = factor(ecog1)))
  p <- p + geom_boxplot(aes(x = ECOG, y = logBTS), alpha = 0.25, shape = 1,
    colour = cbPalette$blue)
  p
```

This code gives us baseline tumour size correlated with ECOG value. While it
doesn't have the exact proportions above and below the median tumour size as the
Joseph et al. paper, it gets close and provides a distribution that looks
similar to that shown in the same paper.

Ideally when used for the study population, these values will be truncated. 5%
of the data being lower than 10 mm and 2.5% being above 200 mm does not feel
particularly useful for a simulation dataset. 

The copula above can be reproduced with existing functions to reduce the need
for designing and validating a new function.

```{r trunc_rnorm}
# First define the demographic data
  btsecog0_mean <- 43.4  # 34.8
  btsecog0_sd <- 35.0  # 29.6
  btsecog1_mean <- 62.2  # 67.7
  btsecog1_sd <- 52.6  # 61.8
  
  bts_range <- c(10, 300)
  
  ecog1_prob <- 0.6126
  
# Sample covariate values from distributions
  ECOG <- rbinom(n, 1, ecog1_prob)
  BTS_ECOG0 <- trunc_rnorm(n = length(ECOG[ECOG == 0]), mean = btsecog0_mean, 
    sd = btsecog0_sd, range = bts_range, log = T)
  BTS_ECOG1 <- trunc_rnorm(n = length(ECOG[ECOG == 1]), mean = btsecog1_mean, 
    sd = btsecog1_sd, range = bts_range, log = T)
  
# Bind covariate data.frame
  cov_df <- data.frame(ECOG)
  cov_df$TUM_0 <- 0
  cov_df$TUM_0[cov_df$ECOG == 0] <- BTS_ECOG0
  cov_df$TUM_0[cov_df$ECOG == 1] <- BTS_ECOG1
  cov_df$ECOGf <- factor(cov_df$ECOG)
  
# Observe the data
  with(cov_df, c(mean(TUM_0), sd(TUM_0), sum(ECOG)/n))
  with(cov_df, c(
    sum(TUM_0[ECOG == 1] > median(TUM_0))/sum(ECOG), 
    sum(TUM_0[ECOG == 0] > median(TUM_0))/sum(1 - ECOG)
  ))
  with(cov_df, c(
    min = min(TUM_0), quantile(TUM_0, prob = 0.05), quantile(TUM_0, prob = 0.25), 
    quantile(TUM_0, prob = 0.5), quantile(TUM_0, prob = 0.75), 
    quantile(TUM_0, prob = 0.95), max = max(TUM_0)
  ))
  
# Plot
  p <- NULL
  p <- ggplot(aes(TUM_0, colour = ECOGf), data = cov_df)
  p <- p + geom_density(size = 1)
  p
  
  p <- NULL
  p <- ggplot(data = cov_df)
  p <- p + geom_boxplot(aes(x = ECOGf, y = log(TUM_0)), alpha = 0.25, shape = 1)
  p
```
