---
title: "Validation of Nivolumab Model"
author: "Jim Hughes"
date: "1 March 2019"
output:
  word_document: default
  html_document: default
---
```{r setup, include=FALSE}
# Load package libraries
  library(plyr)	      # New plyr - required for mrgsolve
  library(dplyr)	    # Split and rearrange data - required for mrgsolve
  library(mrgsolve)	  # Metrum differential equation solver for pharmacometrics
  library(ggplot2)    # Graphical package
  
# Set ggplot2 theme
  theme_bw2 <- theme_set(theme_bw(base_size = 14))
  theme_update(plot.title = element_text(hjust = 0.5))
  
# Set colourblind palette
  cbPalette <- data.frame(
		grey = "#999999",
		orange = "#E69F00",
		skyblue = "#56B4E9",
		green = "#009E73",
		yellow = "#F0E442",
		blue = "#0072B2",
		red = "#D55E00",
		pink = "#CC79A7",
		stringsAsFactors = F
	)  
  myPalette <- with(cbPalette, c(blue, red))
  
# Set project directory
  proj.dir <- "C:/Users/Jim Hughes/Documents/GitRepos/nivo_sim"

# Source processSIMdata function
  nmsim.data <- readRDS(paste0(proj.dir, "/rds/190301_nonmem_process_sim.rds"))
  mrgnm.data <- readRDS(paste0(proj.dir, "/rds/190302_NivoPKPD_Reg.rds"))
  mrgrn.data <- readRDS(paste0(proj.dir, "/rds/190302_NivoPKTS_Reg.rds"))
  # mrgpk.data
```
### Aim
1. Use NONMEM model received from authors to replicate plots from paper
    + Try to replicate points from Figure 3
    + Uses hazard ratio, but at what time?
2. Determine if `mrgsolve` model is coded correctly by validating it against the
NONMEM simulation
    + Try to replicate points from Figure 3
    + Use hazard ratio at time determined above
3. Compare `mrgsolve` version of NONMEM model to the full PKPD model coded in
`mrgsolve`
    + Compare concentration, tumour size and hazard ratio
    
### 1. NONMEM Model
Simulated model using a population of 1000 individuals. Model only has ECOG
as a covariate. This was set to $0$ for all individuals. Individuals were 
simulated with the same seed, giving each patient the same ETA value each time 
the population was simulated.

These patients were based off of the definition of a typical subject as stated
in the manuscript.

> A typical subject (reference) is female, weighing 80 kg, eGFR of 
80 mL/min/1.73 m2 , serum albumin of 4 mg/dL, tumor size of 60 mm, cell 
type/histology of other (i.e., not squamous or non-squamous), performance 
status of 0, tumor type of NSCLC, and ADA assay negative. The reference values 
for continuous covariates were selected to approximate the median values.

The population was simulated with three different dosages from the paper:

* Placebo (0 mg every 2 weeks)
* 20 mg every 2 weeks
* 1280 mg every 2 weeks

This allowed for dosed patients to be matched with their "undosed" self,
allowing the calculation of each individuals relative hazard ratio when taking
nivolumab. The point estimate (mean) of the hazard ratio was calculated to
compare against the figure from the plot.

```{r nmmatch}
# ID represents Dose group
# SIM represents patient ID
# ID == 1 - placebo group
# ID == 2 - 20mg q2w
# ID == 3 - 1280mg q2w
# Want to match the patients from the dose groups with their respective
#   placebo individual
  nmsim.match <- ddply(nmsim.data[nmsim.data$DOSEGRP != 1,], .(DOSEGRP), function(df) {
    cbind(df, 
      data.frame(
        HAZPLAC = nmsim.data$HAZRATE[nmsim.data$DOSEGRP == 1]
      )
    )
  })

# Calculate the hazard ratio
  nmsim.match$HAZRAT <- with(nmsim.match, HAZRATE/HAZPLAC)
  
# Calculate summary statistics
  hazdata <- ddply(nmsim.match, .(DOSEGRP, TIME), function(df) {
    data.frame(
      MEAN = mean(df$HAZRAT, na.rm = T),
      MED = median(df$HAZRAT, na.rm = T),
      CI90LO = quantile(df$HAZRAT, prob = 0.05, na.rm = T),
      CI90HI = quantile(df$HAZRAT, prob = 0.95, na.rm = T)
    )
  })
```

In Figure 3 a dose of 20 mg every 2 weeks has a hazard ratio of ~$0.4$. Therefore
the time corresponding with a hazard ratio of $0.4$ should be the time at which
the hazard ratio was calculated. This can then be validated using the second
population (1280 mg every 2 weeks).

```{r nmhaz}
# Find out what time the hazard ratio is at 0.4
  deltaMEAN <- abs(hazdata$MEAN - 0.4)
  min.deltaMEAN <- min(deltaMEAN, na.rm = T)
  closest <- which(deltaMEAN == min.deltaMEAN)
  hazdata[hazdata$TIME == hazdata$TIME[closest],]
```

This shows that the mean was used to calculate the hazard ratio and that the
time at which the hazard ratio is $0.4$ is 112 days 
(16 weeks). Additionally, it shows the hazard ratio for the 1280 mg every 2 
weeks group is ~$0.1$ which is the same as what is seen in Figure 3 from the 
paper!

### 2. `mrgsolve` Model of NONMEM Model
The same simulations were done for the `mrgsolve` model, using the ETAs from the
NONMEM output. With the hazard ratio of the matched data being observed.

```{r mrgnmhaz, echo = FALSE}
# ID represents Dose group
# SIM represents patient ID
# ID == 1 - placebo group
# ID == 2 - 20mg q2w
# ID == 3 - 1280mg q2w
# Want to match the patients from the dose groups with their respective
#   placebo individual
  mrgnm.match <- ddply(mrgnm.data[mrgnm.data$DOSEGRP != 1,], .(DOSEGRP), function(df) {
    cbind(df, 
      data.frame(
        HAZPLAC = mrgnm.data$HAZRATE[mrgnm.data$DOSEGRP == 1]
      )
    )
  })

# Calculate the hazard ratio
  mrgnm.match$HAZRAT <- with(mrgnm.match, HAZRATE/HAZPLAC)
  
# Calculate summary statistics
  mrgnm.hazr <- ddply(mrgnm.match, .(DOSEGRP, time), function(df) {
    data.frame(
      MEAN = mean(df$HAZRAT, na.rm = T),
      MED = median(df$HAZRAT, na.rm = T),
      CI90LO = quantile(df$HAZRAT, prob = 0.05, na.rm = T),
      CI90HI = quantile(df$HAZRAT, prob = 0.95, na.rm = T),
      MODEL = "Hazard"
    )
  })
  
# Find out what time the hazard ratio is at 0.4
  mrgnm.hazr[mrgnm.hazr$time == 112,]
```

From this we can see that the mrgsolve model is not only able to replicate the
populations simulated in NONMEM, it is also able to replicate the rest of the
points from Figure 3!

### 3. `mrgsolve` Model using PK Model from Manuscript

The same simulations were once again done for the pharmacokinetic model sourced
from the manuscript, using the ETAs from the NONMEM output. The ETA for the
Tmax parameter involved in the calculation of the time-dependent clearance was
sampled from a random distribution.

```{r mrgrnhaz, echo = FALSE}
# ID represents Dose group
# SIM represents patient ID
# ID == 1 - placebo group
# ID == 2 - 20mg q2w
# ID == 3 - 1280mg q2w
# Want to match the patients from the dose groups with their respective
#   placebo individual
  mrgrn.match <- ddply(mrgrn.data[mrgrn.data$DOSEGRP != 1,], .(DOSEGRP), function(df) {
    cbind(df, 
      data.frame(
        HAZPLAC = mrgrn.data$HAZRATE[mrgrn.data$DOSEGRP == 1]
      )
    )
  })

# Calculate the hazard ratio
  mrgrn.match$HAZRAT <- with(mrgrn.match, HAZRATE/HAZPLAC)
  
# Calculate summary statistics
  mrgrn.hazr <- ddply(mrgrn.match, .(DOSEGRP, time), function(df) {
    data.frame(
      MEAN = mean(df$HAZRAT, na.rm = T),
      MED = median(df$HAZRAT, na.rm = T),
      CI90LO = quantile(df$HAZRAT, prob = 0.05, na.rm = T),
      CI90HI = quantile(df$HAZRAT, prob = 0.95, na.rm = T),
      MODEL = "Modified PK"
    )
  })
  
# Find out what time the hazard ratio is at 0.4
  mrgrn.hazr[mrgrn.hazr$time == 112,]
```

This model shows that the hazard rate is higher, particularly at lower doses,
with the rates converging as doses get higher.

### Comparison of the concentrations of the `mrgsolve` models

While it can be seen that the models differ in the hazard rate, visualising
this with plots may provide a better idea of how the predictions of these two
models differ. All comparisons look at doses of 160 and 320 mg every 2 weeks

```{r mrgcomparison, include = FALSE}
# Determine summary statistics for concentration and tumour size
  mrgnm.conc <- ddply(mrgnm.match, .(DOSEGRP, time), function(df) {
    data.frame(
      MEAN = mean(df$IPRED, na.rm = T),
      MED = median(df$IPRED, na.rm = T),
      CI90LO = quantile(df$IPRED, prob = 0.05, na.rm = T),
      CI90HI = quantile(df$IPRED, prob = 0.95, na.rm = T),
      MODEL = "Hazard"
    )
  })

  mrgrn.conc <- ddply(mrgrn.match, .(DOSEGRP, time), function(df) {
    data.frame(
      MEAN = mean(df$IPRED, na.rm = T),
      MED = median(df$IPRED, na.rm = T),
      CI90LO = quantile(df$IPRED, prob = 0.05, na.rm = T),
      CI90HI = quantile(df$IPRED, prob = 0.95, na.rm = T),
      MODEL = "Modified PK"
    )
  })
  
  conc.data <- rbind(mrgnm.conc, mrgrn.conc)
  conc.data$DOSEGRPf <- factor(conc.data$DOSEGRP)
  levels(conc.data$DOSEGRPf) <- c(10, 20, 40, 80, 160, 320, 640, 1280, 2560)

  mrgnm.tums <- ddply(mrgnm.match, .(DOSEGRP, time), function(df) {
    data.frame(
      MEAN = mean(df$TUM, na.rm = T),
      MED = median(df$TUM, na.rm = T),
      CI90LO = quantile(df$TUM, prob = 0.05, na.rm = T),
      CI90HI = quantile(df$TUM, prob = 0.95, na.rm = T),
      MODEL = "Hazard"
    )
  })
  
  mrgrn.tums <- ddply(mrgrn.match, .(DOSEGRP, time), function(df) {
    data.frame(
      MEAN = mean(df$TUM, na.rm = T),
      MED = median(df$TUM, na.rm = T),
      CI90LO = quantile(df$TUM, prob = 0.05, na.rm = T),
      CI90HI = quantile(df$TUM, prob = 0.95, na.rm = T),
      MODEL = "Modified PK"
    )
  })
  
  tums.data <- rbind(mrgnm.tums, mrgrn.tums)
  tums.data$DOSEGRPf <- factor(tums.data$DOSEGRP)
  levels(tums.data$DOSEGRPf) <- c(10, 20, 40, 80, 160, 320, 640, 1280, 2560)
  
  hazr.data <- rbind(mrgnm.hazr, mrgrn.hazr)
  hazr.data$DOSEGRPf <- factor(hazr.data$DOSEGRP)
  levels(hazr.data$DOSEGRPf) <- c(10, 20, 40, 80, 160, 320, 640, 1280, 2560)
  
```

```{r concplotFortnight, echo = FALSE}
# Plot data
  p <- NULL
  p <- ggplot(data = conc.data[conc.data$DOSEGRP %in% c(6, 7), ])
  p <- p + geom_line(aes(x = time, y = MED, colour = MODEL), 
    size = 1)
  p <- p + geom_line(aes(x = time, y = CI90LO, colour = MODEL), 
    linetype = "dashed", size = 1, alpha = 0.5)
  p <- p + geom_line(aes(x = time, y = CI90HI, colour = MODEL), 
    linetype = "dashed", size = 1, alpha = 0.5)
  p <- p + scale_colour_manual(values = myPalette)
  p <- p + facet_wrap(~DOSEGRPf)
  p <- p + ylab("Concentration (mg/L)")
  conc.plot1 <- p + scale_x_continuous("Time (days)", breaks = 0:7*2)
  conc.plot1 <- conc.plot1 + coord_cartesian(xlim = c(0, 14), ylim = c(0, 100))
  conc.plot1
```

This plot shows the median and upper/lower 90% confidence intervals of the
concentrations over the first week. As we can see the 
modified PK model from the manuscript estimates predicts a larger proportion
of elimination over a week. Less drug would relate to a higher hazard than that
seen in the original hazard model.

```{r concplotTwoYears, echo = FALSE}
# Plot data
  conc.plot2 <- p + scale_x_continuous("Time (days)", breaks = 0:4*182)
  conc.plot2
```

Here we have a similar plot but over a two year timespan. It han be seen that
the hazard predicts that drug will accumulate to a great degree, with steady 
state not reached until more than a year has passed. Steady state is reached
far sooner in the modified PK model.

The hazard model that was used for simulations in the manuscript (named the
hazard model in plots) lacks the time-varying clearance mechanism (**separate**
from the time-varying tumour size covariate effect on clearance).

```{r concplotFourMonths, echo = FALSE}
# Plot data
  conc.plot3 <- p + scale_x_continuous("Time (days)", breaks = 0:4*56)
  conc.plot3 <- conc.plot3 + coord_cartesian(xlim = c(0, 224))
  conc.plot3
```

At the time that the hazard model looks at the hazard rate for the manuscript
plots ($t = 112$). As they used the mean of the hazard ratio, and didn't look at
the confidence intervals, it would have been at steady state by that point.

Interestingly they saw how this model could predict the hazard ratio of their
observed data, with the average concentration or minimum concentration from the
first cycle being the most indicative of what the hazard ratio was like. This
may be due to the fact that the predictions are not as far off at this time, as
they would be if you took the average across the entire set of concentrations.

### Comparison of the tumour size of the `mrgsolve` models
Given the concentrations seen above, it is assumed that the tumour size 
predicted by the modified PK model will be higher, due to lower drug
concentrations and therefore, lower efficacy.

```{r tumsplotFortnight, echo = FALSE}
# Plot data
  p <- NULL
  p <- ggplot(data = tums.data[tums.data$DOSEGRP %in% c(6, 7), ])
  p <- p + geom_line(aes(x = time, y = MED, colour = MODEL), 
    size = 1)
  p <- p + geom_line(aes(x = time, y = CI90LO, colour = MODEL), 
    linetype = "dashed", size = 1, alpha = 0.5)
  p <- p + geom_line(aes(x = time, y = CI90HI, colour = MODEL), 
    linetype = "dashed", size = 1, alpha = 0.5)
  p <- p + scale_colour_manual(values = myPalette)
  p <- p + facet_wrap(~DOSEGRPf)
  p <- p + ylab("Tumour Size (mm)")
  tums.plot1 <- p + scale_x_continuous("Time (days)", breaks = 0:7*2)
  tums.plot1 <- tums.plot1 + coord_cartesian(xlim = c(0, 14), ylim = c(25, 75))
  tums.plot1
```

With both models the majority of "typical" patients experience tumour growth 
over the first 14 days. 

```{r tumsplotTwoYears, echo = FALSE}
# Plot data
  tums.plot2 <- p + scale_x_continuous("Time (days)", breaks = 0:4*182)
  tums.plot2
```

Over time patients predicted by the modified PK model have larger tumour sizes
as expected from concentration plots, however there is a large overlap of the
confidence intervals, so they may not be all that different?

### Comparison of the hazard rates of the `mrgsolve` models
While a numerical comparison of hazard rates has been completed, a visual
comparison would also be beneficial.

```{r hazrplotFourMonths, echo = FALSE}
# Plot data
  p <- NULL
  p <- ggplot(data = hazr.data[hazr.data$DOSEGRP %in% c(6, 7), ])
  p <- p + geom_line(aes(x = time, y = MED, colour = MODEL), 
    size = 1)
  p <- p + geom_line(aes(x = time, y = CI90LO, colour = MODEL), 
    linetype = "dashed", size = 1, alpha = 0.5)
  p <- p + geom_line(aes(x = time, y = CI90HI, colour = MODEL), 
    linetype = "dashed", size = 1, alpha = 0.5)
  p <- p + scale_colour_manual(values = myPalette)
  p <- p + facet_wrap(~DOSEGRPf)
  p <- p + ylab("Hazard Ratio")
  hazr.plot1 <- p + scale_x_continuous("Time (days)", breaks = 0:4*56)
  hazr.plot1 <- hazr.plot1 + coord_cartesian(xlim = c(0, 224), ylim = c(0, 1))
  hazr.plot1
```

While the hazard rates differ between the two models, the 90% confidence 
intervals overlap significantly, it is hard to say that they are that different.

To determine whether or not the modified PK is appropriate, it could be 
evaluated in the same fashion as the hazard model; by comparing the hazard rate
to the observed data from the manuscript.

Unfortunately this figure is made using a different simulated dataset, and it
is not clear how they performed the evaluation. Early attempts at replicating
the evaluation have resulted in confidence intervals far wider than those seen
in Figure 4 suggesting inaccurate replication.

