---
title: "Simulation of Troughs from Different Nivolumab Models"
author: "Jim Hughes"
date: "26 March 2019"
output:
  word_document: default
  html_document: default
---
```{r setup, include=FALSE}
# Load package libraries
  library(dplyr)	    # Split and rearrange data - required for mrgsolve
  library(mrgsolve)	  # Metrum differential equation solver for pharmacometrics
  library(ggplot2)    # Graphical package
  library(knitr)      # kable
  
# Set ggplot2 theme
  theme_bw2 <- theme_set(theme_bw(base_size = 14))
  theme_update(plot.title = element_text(hjust = 0.5))
  
# Set colourblind palette
  cbPalette <- data.frame(
		grey = "#999999",
		orange = "#E69F00",
		skyblue = "#56B4E9",
		green = "#009E73",
		yellow = "#F0E442",
		blue = "#0072B2",
		red = "#D55E00",
		pink = "#CC79A7",
		stringsAsFactors = F
	)  
  myPalette <- with(cbPalette, c(blue, red))
  
# Set project directory
  proj.dir <- "C:/Users/hugjh001/Documents/nivo_sim"

# Source processSIMdata function
  PKPD_240 <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKPD_Reg_240.rds"))
  PKPD_240nt <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKPD_Reg_240_notum.rds"))
  PKPD_240pp <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKPD_Reg_240_PKPDpop.rds"))
  PKPD_80 <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKPD_Reg_80.rds"))
  PKTSa_240 <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKTS_Reg_240.rds"))
  PKTSa_240nt <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKTS_Reg_240_notum.rds"))
  PKTSa_240pp <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKTS_Reg_240_PKPDpop.rds"))
  PKTSa_80 <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKTS_Reg_80.rds"))
  PKTSb_240 <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKTSalt_Reg_240.rds"))
  PKTSb_240nt <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKTSalt_Reg_240_notum.rds"))
  PKTSb_240pp <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKTSalt_Reg_240_PKPDpop.rds"))
  PKTSb_80 <- readRDS(paste0(proj.dir, "/rds/190326_NivoPKTSalt_Reg_80.rds"))
  
# Create geometric mean function
  gm_mean = function(x, na.rm=TRUE){
    exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
  }
```
### Aim
Determine what are the average troughs for each of the nivolumab models and
whether they match the data given in the FDA application for nivolumab.

### FDA Data
The FDA submission for nivolumab included the mean troughs for patients at the
beginning of the ninth cycle (day 112) Link to the pdf* and table below:

```{r FDA}
  kable(data.frame(
    Dose_mgkg = c(0.1, 0.3, 1, 3, 10),
    Patients_n = c(7, 5, 35, 21, 44),
    GeoMean_Trough_ugmL = c(2.5, 6.9, 19, 57, 188.8)
  ))
```

> https://www.accessdata.fda.gov/drugsatfda_docs/nda/2014/125554Orig1s000ClinPharmR.pdf 

> *Table present on page 16 of the .pdf

### Models
Models are classifed as:

* NONMEM Model (model used for simulation by in Liu et al.)
* Modified-PK Model (developed PK model from Liu et al. with tumour growth from
their simulations)
* Alternative PK Model (Modified-PK Model without the additional 
tumour-independent time-varying term on clearance)

#### Melanoma Patients
Here are the troughs for the melanoma population (to be used in our simulation
study and referred to in `nivo_population.rmd`)

```{r models, echo = F}
  kable(data.frame(
    Dose_mgkg = c(1, 3),
    NONMEM_Model = c(
      gm_mean(dplyr::filter(PKPD_80, time == 112)$DV),
      gm_mean(dplyr::filter(PKPD_240, time == 112)$DV)
    ),
    ModPK_Model = c(
      gm_mean(dplyr::filter(PKTSa_80, time == 112)$DV),
      gm_mean(dplyr::filter(PKTSa_240, time == 112)$DV)
    ),
    AltPK_Model = c(
      gm_mean(dplyr::filter(PKTSb_80, time == 112)$DV),
      gm_mean(dplyr::filter(PKTSb_240, time == 112)$DV)
    )
  ))
```

#### Healthy Individuals
To determine the impact of tumour size on the troughs in these models, each
model was simulated using "healthy individuals" (intitial tumour size = 0.1 mm).

```{r models2, echo = F}
  kable(data.frame(
    Dose_mgkg = c(3),
    NONMEM_Model = c(
      gm_mean(dplyr::filter(PKPD_240nt, time == 112)$DV)
    ),
    ModPK_Model = c(
      gm_mean(dplyr::filter(PKTSa_240nt, time == 112)$DV)
    ),
    AltPK_Model = c(
      gm_mean(dplyr::filter(PKTSb_240nt, time == 112)$DV)
    )
  ))
```

Tumour size has a large impact on the models. It is possible that the covariate
tumour models were fit to patient data with larger tumours than those seen in 
the FDA submission, resulting in different trough concentrations at 3mg/kg.

#### Liu et al. Simulation Population
To determine the impact of the population on the troughs, each model was 
simulated using the standard individual used in the NONMEM model simulation.

```{r models3, echo = F}
  kable(data.frame(
    Dose_mgkg = c(3),
    NONMEM_Model = c(
      gm_mean(dplyr::filter(PKPD_240pp, time == 112)$DV)
    ),
    ModPK_Model = c(
      gm_mean(dplyr::filter(PKTSa_240pp, time == 112)$DV)
    ),
    AltPK_Model = c(
      gm_mean(dplyr::filter(PKTSb_240pp, time == 112)$DV)
    )
  ))
```

This shows that the NONMEM model is possibly the same model as the altPK model,
except set up in such a way that all the covariates are gone.

#### Which model?
The modified PK model is likely still the best, as it shows the true variability
in patient clearance due to time-varying clearance due to disease outcome, which
is missing from the NONMEM and alternative PK models.

### Trough Plot
```{r plot, echo = F, fig.asp = 1.5}
# Bring data together
  PKPD_240$MOD <- "NONMEM"
  PKPD_240nt$MOD <- "NONMEM"
  PKPD_240pp$MOD <- "NONMEM"
  PKTSa_240$MOD <- "Mod_PK"
  PKTSa_240nt$MOD <- "Mod_PK"
  PKTSa_240pp$MOD <- "Mod_PK"
  PKTSb_240$MOD <- "Alt_PK"
  PKTSb_240nt$MOD <- "Alt_PK"
  PKTSb_240pp$MOD <- "Alt_PK"
  
  PKPD_240$SIM <- "Melanoma Patients"
  PKTSa_240$SIM <- "Melanoma Patients"
  PKTSb_240$SIM <- "Melanoma Patients"
  PKPD_240nt$SIM <- "Healthy Individuals"
  PKTSa_240nt$SIM <- "Healthy Individuals"
  PKTSb_240nt$SIM <- "Healthy Individuals"
  PKPD_240pp$SIM <- "Liu et al. Sim. Pop."
  PKTSa_240pp$SIM <- "Liu et al. Sim. Pop."
  PKTSb_240pp$SIM <- "Liu et al. Sim. Pop."
  
# There is almost definitely a better way of doing this, but here we go
  plot_df <- tibble::tibble(
      ID_data = 1:9,
      data = list(
        PKPD_240,
        PKPD_240nt,
        PKPD_240pp,
        PKTSa_240,
        PKTSa_240nt,
        PKTSa_240pp,
        PKTSb_240,
        PKTSb_240nt,
        PKTSb_240pp
      )
    ) %>%
    dplyr::mutate(data = 
      purrr::map(data, function(data) {
        dplyr::select(data, ID, time, DV, MOD, SIM)
      })
    ) %>%
    tidyr::unnest()
  
# Make plots
  p <- NULL
  p <- ggplot(data = plot_df)
  p <- p + stat_summary(aes(x = time, y = DV, group = MOD, colour = MOD),
    fun.y = gm_mean, geom = "line", size = 1)
  p <- p + geom_hline(yintercept = 57, linetype = "dashed")
  p <- p + geom_vline(xintercept = 112, linetype = "dashed")
  p <- p + scale_x_continuous("Time (days)", breaks = 0:3*112)
  p <- p + scale_y_continuous("Concentration (mg/mL)",  
    breaks = c(3, 10, 30, 100, 300))
  p <- p + coord_trans(y = "log10", limy = c(2.5, 112))
  p <- p + guides(colour = guide_legend(title = "Model"))
  p <- p + facet_wrap(~SIM, ncol = 1)
  p
```